Combinatorial Proofs and Decomposition Theorems for First-order Logic\vspace{-3ex}

### 2021-01-25 main.tex
typo
todo
backing up changes to section 1

\begin{document}
\vlnosmallleftlabels

\bstctlcite{refs:BSTcontrol}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Combinatorial Proofs and Decomposition Theorems for First-order Logic}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
  We uncover a close relationship between combinatorial and
  syntactic proofs for first-order logic. Whereas syntactic proofs are
  formalized in some deductive proof system based on inference rules,
  a combinatorial proof is a ``syntax-free'' presentation of a proof
  that is independent from any set of inference rules. We show that
  the two proof representations are related via a deep inference
  decomposition theorem that establishes a new kind of normal form
  of syntactic proofs. This yields (a) a simple proof
  of the soundness and completeness of first-order combinatorial
  proofs, and (b) a full completeness theorem: every combinatorial
  proof is the image of a syntactic proof.
\end{abstract}

@@

The main reason for this problem is that formal proofs, as they are
usually studied in logic, are inextricably tied to the syntactic
(inference rule based) proof system in which they are carried out. And
it is difficult to compare two proofs that are produced within two
different syntactic proof systems, based on different sets of inference rules.

\begin{figure}
  $$
  \scalebox{.9}{$
\vlderivation{
  \vlin{\vlor}{}{\sqn{((\cneg{p} \vlor q) \vlan \cneg{p}) \vlor p}}{
    \vlin{\conr}{}{\sqn{(\cneg{p} \vlor q) \vlan \cneg{p}, p}}{
      \vliin{\vlan}{}{\sqn{(\cneg{p} \vlor q) \vlan \cneg{p}, p, p}}{
        \vlin{\vlor}{}{\sqn{\cneg{p} \vlor q, p}}{
          \vlin{\weakr}{}{\sqn{\cneg{p}, q, p}}{
            \vlin{\axr}{}{\sqn{\cneg{p}, p}}{
              \vlhy{}}}}}{
        \vlin{\axr}{}{\sqn{\cneg{p}, p}}{
          \vlhy{}}}}}}
\hskip3em
\vlderivation{
  \vlin{\acD}{}{((\cneg{p} \vlor q) \vlan \cneg{p}) \vlor p}{
    \vlin{\fequ}{}{((\cneg{p} \vlor q) \vlan \cneg{p}) \vlor (p \vlor p)}{
      \vlin{\sw}{}{(\cneg{p} \vlan (\cneg{p} \vlor q)) \vlor p) \vlor p}{
        \vlin{\fequ}{}{(\cneg{p} \vlan ((\cneg{p} \vlor q) \vlor p)) \vlor p)}{
          \vlin{\sw}{}{(p \vlor (\cneg{p} \vlor q)) \vlan \cneg{p}) \vlor p}{
            \vlin{\aiD}{}{(p \vlor (\cneg{p} \vlor q)) \vlan (\cneg{p} \vlor p)}{
              \vlin{\tttD}{}{(p \vlor (\cneg{p} \vlor q)) \vlan \ttt}{
                \vlin{\fequ}{}{p \vlor (\cneg{p} \vlor q)}{
                  \vlin{\wrD}{}{(p \vlor \cneg{p}) \vlor q}{
                    \vlin{\aiD}{}{p \vlor \cneg{p}}{
                      \vlhy{\ttt}}}}}}}}}}}}
$}
  $$
  \caption{Two syntactic proofs of the same formula, in in the sequent calculus (left) and in the calculus of structures (right)}
  \label{fig:exa-synt}
\end{figure}

\lutz{an example here?}

This is where \emph{combinatorial proofs} come in. They have been
introduced by Hughes~\cite{hughes:pws} for classical propositional
logic as a ``syntax-free'' notion of proof, and as a potential solution
to Hilbert's 24th problem~\cite{hughes:invar} (see
also~\cite{str:hilbert:24}). The basic idea is to abstract away from
the syntax of the inference rules used in the proof and consider the
proof as a combinatorial object, more precisely as a special kind of
graph homomorphism. For example, a propositional combinatorial
proof of Peirce's law $\peirceimpliesformula$ is depicted below-left,
a homomorphism from a coloured graph to a graph labelled with
propositional variables.
%
\begin{center}\vspace{.5ex}\begin{tikzpicture}\newcommand\radius{2.3}
  \begin{scope}[shift={(-\radius,0)}]{\peircecp}\end{scope}
  \begin{scope}[shift={(\radius,0)}]{\drinkerfibcoloured}\end{scope}
\end{tikzpicture}\vspace{0ex}\end{center}
%

\begin{equation*}
  \scalebox{.7}{$
\vlderivation{
  \vlin{\conr}{}{\sqn{\exists x.(\cneg{p}x \vlor (\forall y.py))}}{
    \vlin{\exists}{}{\sqn{\exists x.(\cneg{p}x \vlor (\forall y.py)),\exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
      \vlin{\vlor}{}{\sqn{\cneg{p}w \vlor (\forall y.py), \exists
x.(\cneg{p}x \vlor (\forall y.p(y)))}}{
        \vlin{\forall}{}{\sqn{\cneg{p}w, \forall y.py, \exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
          \vlin{\exists}{}{\sqn{\cneg{p}w, pz, \exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
            \vlin{\vlor}{}{\sqn{\cneg{p}w, pz, \cneg{p}z \vlor (\forall y.py))}}{
              \vlin{\weakr}{}{\sqn{\cneg{p}w, pz, \cneg{p}z, (\forall y.py))}}{
                \vlin{\weakr}{}{\sqn{\cneg{p}w, pz, \cneg{p}z}}{
                \vlin{\axr}{}{\sqn{pz, \cneg{p}z}}{
                  \vlhy{}}}}}}}}}}}
%\qquad
\vlderivation{
  \vlin{\acD}{}{\exists x.(\cneg{p}x \vlor (\forall y.py))}{
    \vlin{\mfaD}{}{\exists x.(\cneg{p}x \vlor (\forall y.(py \vlor py)))}{
      \vlin{\acD}{}{\exists x.(\cneg{p}x \vlor ((\forall y.py) \vlor (\forall
y.py)))}{
        \vlin{\fequ}{}{\exists x.((\cneg{p}x \vlor \cneg{p}x) \vlor ((\forall
y.py) \vlor (\forall y.py)))}{
          \vlin{\mexD}{}{\exists x.((\cneg{p}x \vlor (\forall y.py)) \vlor
(\cneg{p}x \vlor (\forall y.py)))}{
            \vlin{\exists}{}{(\exists x.(\cneg{p}x \vlor (\forall y.py))) \vlor
(\exists x.(\cneg{p}x \vlor (\forall y.py)))}{
              \vlin{\fequ}{}{(\cneg{p}w \vlor (\forall y.py)) \vlor (\exists
x.(\cneg{p}x \vlor (\forall y.py)))}{
                \vlin{\exists}{}{\forall y.((\cneg{p}w \vlor py) \vlor (\exists
x.(\cneg{p}x \vlor (\forall y.py))))}{
                  \vlin{\fequ}{}{\forall y.((\cneg{p}w \vlor py) \vlor
(\cneg{p}y \vlor (\forall y.py)))}{
                    \vlin{\wD}{}{\forall y.((py \vlor \cneg{p}y) \vlor
(\cneg{p}w \vlor (\forall y.py)))}{
                      \vlin{\aiD}{}{\forall y.(py \vlor \cneg{p}y)}{
                        \vlin{\forall}{}{\forall y.\ttt}{
                          \vlhy{\ttt}}}}}}}}}}}}}}
$}
\end{equation*}

\juihsuan{LK1 and KS1 proofs of Peirce's law and LK1 proof of drinker formula}
\lutz{can you also do the KS1 for the drinker?}
\juihsuan{done}

Several authors have illustrated how syntactic proofs in various
proof systems can be translated to propositional combinatorial proofs:
for sequent proofs in~\cite{hughes:invar}, for deep inference proofs
in~\cite{str:fscd17}, for Frege systems in~\cite{str:RR-9048}, and for
tableaux systems and resolution in~\cite{acc:str:18}. This enables a
natural definition of proof identity for propositional logic: two
proofs are \emph{the same}, if they are mapped to the same
combinatorial proof.

@@

In this paper we advance the idea that combinatorial proofs can provide
a notion of proof identity for first-order logic. \emph{First-order
combinatorial proofs} were introduced by Hughes in~\cite{hughes:fopws}.
For example, a first-order combinatorial proof of Smullyan's 
``drinker paradox'' $\drinkerformula$ is shown above-right,
a homomorphism from a partially coloured graph to a labelled graph.
However, even though Hughes proves soundness and completeness, the 
proof is highly unsatisfactory: (1) the soundness argument is extremely
long, intricate and cumbersome, and (2) the completeness proof does not 
allow a syntactic proof to be read back from a combinatorial proof, i.e.,
completeness is not \emph{full}~\cite{abramsky:jagadeesan:94}.\todo{cite something for full comp?}\lutz{is that enough?}
A fundamental problem is that not all combinatorial
proofs can be obtained as translations of sequent calculus proofs.

In this paper we solve this issue by moving to a deep inference
system. More precisely, we introduce a new proof system, called
$\FOKS$, for first-order logic, that (a) reflects every combinatorial
proof, i.e., there is a surjective mapping from proofs in $\FOKS$ to
combinatorial proofs, and (b) yields far simpler proofs of
soundness and completeness of combinatorial proofs, and (c) admits
new decomposition theorems establishing a precise correspondence
between certain syntactic inference rules and certain combinatorial
notions.

@@

Combinatorial proofs and deep inference can be seen as opposite ends
of a spectrum: whereas deep inference allows for a very fine
granularity of inference rules---one inference rule in a standard
formalism, like sequent calculus or semantic tableaux, is usually
simulated by a sequence of different deep inference rules---have
combinatorial proofs completely abolished the concept of inference
rule. Nonetheless, there is a close relationship between the two,
realized through a decomposition theorem, as we establish it in this
paper.

\subsection{Pictures that could be used elsewhere in the paper}

Condensed combinatorial proof of drinker formula(s):
%
\begin{center}\vspace{3ex}\begin{math}
\begin{tikzpicture}
\drinkerInline
\end{tikzpicture}
\hspace{10ex}
\begin{tikzpicture}
\veedrinkerInline
\end{tikzpicture}\end{math}\end{center}

@@

Illustrating why we need to collapse twins during contraction, to preserve the target as a fograph:
%
\twincollapsepic
%
Aligning both $\singleton{}\mkern4mu\singleton[red]{}$ and
$\singleton{}\mkern4mu\singleton[blue]{}$ over
a single copy of $\forall y\mkern2mu\py$
results in two uncoloured vertices
\raisebox{2pt}[0ex][0ex]{$\column{\singleton{}\\[-1.6ex]\singleton{}}$\hspace{-2pt}}
over $\forall y$.
%
The cover therefore fails to be a fograph: both uncoloured vertices are implicitly labelled with $y$ and
so are outer $y$-binders in the scope of each other.
%
The correct operation is shown above-right,
in which the troublesome pair is collapsed to a single uncoloured vertex $\singleton{}$ over $\forall y$.

@@

In this paper we use the sequent calculus $\FOLK$, shown in
Figure~\ref{fig:LK1}, which is a one-sided variant of Gentzen's
original calculus~\cite{gentzen:35:I} for first-order logic. To
simplify some technicalities later in this paper, we include
the $\mixr$-rule.
%
\begin{thm}
  $\FOLK$ is sound and complete for first-order logic.
\end{thm}
%
For a proof we refer to reader to any standard textbook, e.g.~\cite{TS:00}.

@@

\begin{example}
  Below is the fograph of drinker formula $\drinkerformula=\veedrinkerformula$:
%
\begin{center}\vspace{2ex}\begin{tikzpicture}
\drinkerbase
\end{tikzpicture}\vspace{1ex}\end{center}
%
%
\end{example}

@@

The obvious next step in our research is to investigate proof
composition and normalisation of first-order combinatorial
proofs. Even in the propositional setting, the normalization of
combinatorial proofs is underdevelopped. There exist two different
procedures for cut elimination~\cite{hughes:invar,str:fscd17}, but
both have their insufficiencies, and there is no general theory.

@@

>>>>

\begin{document}
\vlnosmallleftlabels%
\bstctlcite{refs:BSTcontrol}%
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Combinatorial Proofs and Decomposition Theorems for First-order Logic\vspace{-3ex}}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
  We uncover a close relationship between combinatorial and
  syntactic proofs for first-order logic. Whereas syntactic proofs are
  formalized in a deductive proof system based on inference rules,
  a combinatorial proof is a syntax-fre presentation of a proof
  that is independent from any set of inference rules. We show that
  the two proof representations are related via a deep inference
  decomposition theorem that establishes a new kind of normal form
  for syntactic proofs. This yields (a) a simple proof
  of soundness and completeness for first-order combinatorial
  proofs, and (b) a full completeness theorem: every combinatorial
  proof is the image of a syntactic proof.
\end{abstract}

@@

The main reason for this problem is that formal proofs, as they are
usually studied in logic, are inextricably tied to the syntactic
(inference rule based) proof system in which they are carried out. And
it is difficult to compare two proofs that are produced within two
different syntactic proof systems, based on different sets of inference rules.
%
\newcommand\lkpeirceproof{
  \vlin{\vlor}{}{\sqn{\peirceformula}}{
    \vlin{\conr}{}{\sqn{(\cneg{p} \vlor q) \vlan \cneg{p}, p}}{
      \vliin{\vlan}{}{\sqn{(\cneg{p} \vlor q) \vlan \cneg{p}, p, p}}{
        \vlin{\vlor}{}{\sqn{\cneg{p} \vlor q, p}}{
          \vlin{\weakr}{}{\sqn{\cneg{p}, q, p}}{
            \vlin{\axr}{}{\sqn{\cneg{p}, p}}{
              \vlhy{}}}}}{
        \vlin{\axr}{}{\sqn{\cneg{p}, p}}{
          \vlhy{}}}}}}
\newcommand\dipeirceproof{
  \vlin{\acD}{}{\peirceformula}{
    \vlin{\fequ}{}{((\cneg{p} \vlor q) \vlan \cneg{p}) \vlor (p \vlor p)}{
      \vlin{\sw}{}{(\cneg{p} \vlan (\cneg{p} \vlor q)) \vlor p) \vlor p}{
        \vlin{\fequ}{}{(\cneg{p} \vlan ((\cneg{p} \vlor q) \vlor p)) \vlor p)}{
          \vlin{\sw}{}{(p \vlor (\cneg{p} \vlor q)) \vlan \cneg{p}) \vlor p}{
            \vlin{\aiD}{}{(p \vlor (\cneg{p} \vlor q)) \vlan (\cneg{p} \vlor p)}{
              \vlin{\tttD}{}{(p \vlor (\cneg{p} \vlor q)) \vlan \ttt}{
                \vlin{\fequ}{}{p \vlor (\cneg{p} \vlor q)}{
                  \vlin{\wrD}{}{(p \vlor \cneg{p}) \vlor q}{
                    \vlin{\aiD}{}{p \vlor \cneg{p}}{
                      \vlhy{\ttt}}}}}}}}}}}}
\newcommand\lkdrinkerproof{
  \vlin{\conr}{}{\sqn{\exists x.(\cneg{p}x \vlor (\forall y.py))}}{
    \vlin{\exists}{}{\sqn{\exists x.(\cneg{p}x \vlor (\forall y.py)),\exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
      \vlin{\vlor}{}{\sqn{\cneg{p}w \vlor (\forall y.py), \exists
x.(\cneg{p}x \vlor (\forall y.p(y)))}}{
        \vlin{\forall}{}{\sqn{\cneg{p}w, \forall y.py, \exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
          \vlin{\exists}{}{\sqn{\cneg{p}w, pz, \exists
x.(\cneg{p}x \vlor (\forall y.py))}}{
            \vlin{\vlor}{}{\sqn{\cneg{p}w, pz, \cneg{p}z \vlor (\forall y.py))}}{
              \vlin{\weakr}{}{\sqn{\cneg{p}w, pz, \cneg{p}z, (\forall y.py))}}{
                \vlin{\weakr}{}{\sqn{\cneg{p}w, pz, \cneg{p}z}}{
                \vlin{\axr}{}{\sqn{pz, \cneg{p}z}}{
                  \vlhy{}}}}}}}}}}}
\newcommand\ksonedrinkerproof{
  \vlin{\acD}{}{\exists x.(\cneg{p}x \vlor (\forall y.py))}{
    \vlin{\mfaD}{}{\exists x.(\cneg{p}x \vlor (\forall y.(py \vlor py)))}{
      \vlin{\acD}{}{\exists x.(\cneg{p}x \vlor ((\forall y.py) \vlor (\forall y.py)))}{
        \vlin{\fequ}{}{\exists x.((\cneg{p}x \vlor \cneg{p}x) \vlor ((\forall y.py) \vlor (\forall y.py)))}{
          \vlin{\mexD}{}{\exists x.((\cneg{p}x \vlor (\forall y.py)) \vlor (\cneg{p}x \vlor (\forall y.py)))}{
            \vlin{\exists}{}{(\exists x.(\cneg{p}x \vlor (\forall y.py))) \vlor (\exists x.(\cneg{p}x \vlor (\forall y.py)))}{
              \vlin{\fequ}{}{(\cneg{p}w \vlor (\forall y.py)) \vlor (\exists x.(\cneg{p}x \vlor (\forall y.py)))}{
                \vlin{\exists}{}{\forall y.((\cneg{p}w \vlor py) \vlor (\exists x.(\cneg{p}x \vlor (\forall y.py))))}{
                  \vlin{\fequ}{}{\forall y.((\cneg{p}w \vlor py) \vlor (\cneg{p}y \vlor (\forall y.py)))}{
                    \vlin{\wD}{}{\forall y.((py \vlor \cneg{p}y) \vlor (\cneg{p}w \vlor (\forall y.py)))}{
                      \vlin{\aiD}{}{\forall y.(py \vlor \cneg{p}y)}{
                        \vlin{\forall}{}{\forall y.\ttt}{
                          \vlhy{\ttt}}}}}}}}}}}}}}
\begin{figure}\vspace{-2ex}\begin{center}\hspace*{-0ex}\begin{math}
\newcommand\peircescale{.75}
\newcommand\drinkerscale{\peircescale}
\newcommand\cpscale{.9}
\begin{array}{@{}c@{\hspace{4ex}}c@{}}
{}\scalebox{\peircescale}{$\lkpeirceproof$}
&
{}\scalebox{\drinkerscale}{$\lkdrinkerproof$}
\\[13ex]
\downarrow
&
\downarrow
\\[1ex]
\vctr{\scalebox{\cpscale}{\peircecppic}}
&
\vctr{\raisebox{-2ex}[16.5ex]{\scalebox{\cpscale}{\drinkerfibcolouredpic}}}
\\[10ex]
\uparrow
&
\uparrow
\\[2ex]
{}\scalebox{\peircescale}{$\dipeirceproof$}
&
{}\scalebox{\drinkerscale}{$\ksonedrinkerproof$}
%
\end{array}
\end{math}\hspace*{-8ex}\end{center}\caption{Left:
syntactic proofs in sequent calculus (above)
and the calculus of structures (below)
which translate to the same propositional combinatorial proof (centre).
%
Right:
syntactic proofs 
in sequent calculus (above)
and the new calculus $\FOKS$ introduced in this paper (below),
which translate to the same first-order combinatorial proof (centre).%
% two syntactic proofs of 
% $\protect\peirceformula$,
% in the sequent calculus (above)
% and in the calculus of structures (below),
% which translate to the same propositional combinatorial proof (centre).
% %
% Right:
% two syntactic proofs of 
% $\protect\veedrinkerformula$,
% in the sequent calculus (above)
% and the new calculus $\FOKS$ introduced in this paper (below),
% which translate to the same first-order combinatorial proof (centre).%
}\label{fig:translate}\vspace{-4ex}\end{figure}%

This is where \emph{combinatorial proofs} come in. They were
introduced by Hughes~\cite{hughes:pws} for classical propositional
logic as a syntax-free notion of proof, and as a potential solution
to Hilbert's 24th problem~\cite{hughes:invar} (see
also~\cite{str:hilbert:24}). The basic idea is to abstract away from
the syntax of the inference rules used in inductively-generated proofs
and consider the proof as a combinatorial object, more precisely
as a special kind of
graph homomorphism. For example, a propositional combinatorial
proof of Peirce's law $\peirceimpliesformula=\peirceformula$ is shown mid-left in
Fig.\,\ref{fig:translate},
%depicted below-left,
a homomorphism from a coloured graph to a graph labelled with
propositional variables.
%
% \begin{center}\vspace{.5ex}\begin{tikzpicture}\newcommand\radius{2.3}
%   \begin{scope}[shift={(-\radius,0)}]{\peircecp}\end{scope}
%   \begin{scope}[shift={(\radius,0)}]{\drinkerfibcoloured}\end{scope}
% \end{tikzpicture}\vspace{0ex}\end{center}
% %
% 
% \begin{equation*}
%   \scalebox{.7}{$\lkdrinkerproof
% %\qquad
% \ksonedrinkerproof
% $}
% \end{equation*}
% 
% \juihsuan{LK1 and KS1 proofs of Peirce's law and LK1 proof of drinker formula}
% \lutz{can you also do the KS1 for the drinker?}
% \juihsuan{done}

Several authors have illustrated how syntactic proofs in various
proof systems can be translated to propositional combinatorial proofs:
for sequent proofs in~\cite{hughes:invar}, for deep inference proofs
in~\cite{str:fscd17}, for Frege systems in~\cite{str:RR-9048}, and for
tableaux systems and resolution in~\cite{acc:str:18}. This enables a
natural definition of proof identity for propositional logic: two
proofs are \emph{the same}, if they are mapped to the same
combinatorial proof.
%
For example, the left side of Fig.\,\ref{fig:translate} translates
syntactic proofs from sequent calculus and the calculus of structures into
the same combinatorial proofs, witnessing that the two syntactic proofs,
from different systems, are \emph{the same}.

@@

In this paper we advance the idea that combinatorial proofs can provide
a notion of proof identity for first-order logic. \emph{First-order
combinatorial proofs} were introduced by Hughes in~\cite{hughes:fopws}.
For example, a first-order combinatorial proof of Smullyan's 
``drinker paradox'' $\drinkerformula=\veedrinkerformula$ is shown on the
right of Fig.\,\ref{fig:translate},
a homomorphism from a partially coloured graph to a labelled graph.
However, even though Hughes proves soundness and completeness, the 
proof is highly unsatisfactory: (1) the soundness argument is extremely
long, intricate and cumbersome, and (2) the completeness proof does not 
allow a syntactic proof to be read back from a combinatorial proof, i.e.,
completeness is not
\emph{sequentializable} \cite{girard:87} or
\emph{full}~\cite{abramsky:jagadeesan:94}.\dominic{ok to say sequentializable too?}
%
A fundamental problem is that not all combinatorial
proofs can be obtained as translations of sequent calculus proofs.

In this paper we solve these issues by moving to a deep inference
system. More precisely, we introduce a new proof system, called
$\FOKS$, for first-order logic, that (a) reflects every combinatorial
proof, i.e., there is a surjection from $\FOKS$ proofs to
combinatorial proofs, (b) yields far simpler proofs of
soundness and completeness of combinatorial proofs, and (c) admits
new decomposition theorems establishing a precise correspondence
between certain syntactic inference rules and certain combinatorial
notions.
%
The right side of Fig.\,\ref{fig:translate}
illustrates the surjection in (a), and since the syntactic proofs 
of the two systems both translate the same combinatorial proof,
they can be considered \emph{the same}.

@@

Combinatorial proofs and deep inference can be seen as opposite ends
of a spectrum: whereas deep inference allows for a very fine
granularity of inference rules---one inference rule in a standard
formalism, like sequent calculus or semantic tableaux, is usually
simulated by a sequence of different deep inference rules---
combinatorial proofs have completely abolished the concept of inference
rule. And yet, there is a close relationship between the two,
realized through a decomposition theorem, as we establish in this
paper.

% \subsection{Pictures that could be used elsewhere in the paper}
% 
% 
% Condensed combinatorial proof of drinker formula(s):
% %
% \begin{center}\vspace{3ex}\begin{math}
% \begin{tikzpicture}
% \drinkerInline
% \end{tikzpicture}
% \hspace{10ex}
% \begin{tikzpicture}
% \veedrinkerInline
% \end{tikzpicture}\end{math}\end{center}

@@

% Illustrating why we need to collapse twins during contraction, to preserve the target as a fograph:
% %
% \twincollapsepic
% %
% Aligning both $\singleton{}\mkern4mu\singleton[red]{}$ and
% $\singleton{}\mkern4mu\singleton[blue]{}$ over
% a single copy of $\forall y\mkern2mu\py$
% results in two uncoloured vertices
% \raisebox{2pt}[0ex][0ex]{$\column{\singleton{}\\[-1.6ex]\singleton{}}$\hspace{-2pt}}
% over $\forall y$.
% %
% The cover therefore fails to be a fograph: both uncoloured vertices are implicitly labelled with $y$ and
% so are outer $y$-binders in the scope of each other.
% %
% The correct operation is shown above-right,
% in which the troublesome pair is collapsed to a single uncoloured vertex $\singleton{}$ over $\forall y$.

@@

In this paper we use the sequent calculus $\FOLK$, shown in
Figure~\ref{fig:LK1}, which is a one-sided variant of Gentzen's
original calculus~\cite{gentzen:35:I} for first-order logic. To
simplify some technicalities later in this paper, we include
the $\mixr$-rule.
%
\begin{thm}
  $\FOLK$ is sound and complete for first-order logic.
\end{thm}
%
\noindent For a proof, see any standard textbook, e.g.~\cite{TS:00}.

@@

\begin{example}
  Below is the fograph of the drinker formula $\drinkerformula=\veedrinkerformula$:
%
\begin{center}\vspace{2ex}\begin{tikzpicture}
\drinkerbase
\end{tikzpicture}\vspace{1ex}\end{center}
%
%
\end{example}

@@

The obvious next step in our research is to investigate proof
composition and normalisation of first-order combinatorial
proofs. Even in the propositional setting, the normalization of
combinatorial proofs is underdeveloped. There exist two different
procedures for cut elimination~\cite{hughes:invar,str:fscd17}, but
both have their insufficiencies, and there is no general theory.

@@

\todo{mention CERES}
